Is structure based drug design ready for selectivity optimization?

### 2018-11-14 manuscript.tex
Updates from ShareLaTeX

\paragraph{Free energy calculation errors are correlated, accelerating selectivity optimization}
To quantify the correlation coefficient ($
\rho$) of the errors in our calculations, we built a Bayesian graphical model, as described in the methods section. Briefly, we modeled the absolute free energy ($G$) of each ligand in each phase (complex and solvent) as in equation~\ref{eq4}. The model was chained to the FEP+ calculations by providing the $\Delta G^{calc}_{phase,ij,target}$ as observed data, as in equation~\ref{eq6}. As in equation, the experimental data was modeled as a normal distribution centered around the true free energy of binding ($\Delta G^{true}_{i,target}$) corrupted by experimental error, which is assumed to be 0.3 kcal/mol from previous work done to quantify the uncertainty in publicly available data~\citep{BROWN2009420}. The reported IC50 values from each dataset were treated as data observations (Equation~\ref{eq12}) and the $\Delta G^{true}_{i,target}$ was assigned a weak normal prior (Equation~\ref{eq13}). The correlation coefficient was calculated for each sample according to equation~\ref{eq9}. 
The correlation coefficient $\rho$ for the CDK2/ERK2 calculations was quantified to be $0.5^{0.33}_{-0.23}$, indicating that the errors are largely uncorrelated between ERK2 and CDK2 (Figure~\ref{fig:figure-5}A, right). The joint marginal distribution of the error ($\epsilon$) for each target is symmetric, which is expected for cases in which $\rho$ is 0 (Supp. Figure~\ref{fig:sup-figure-2}). Despite the weak correlation in errors, the high per target accuracy of these calculations should have a 2-3x speed up for 1 log unit selectivity optimization, and a 20-30x speed up for 2 log unit selectivity optimization (Figure~\ref{fig:figure-5}A, right). 
The CDK2/CDK9 calculations show strong evidence of correlation, with a correlation coefficient of $0.70^{0.82}_{0.57}$ (Figure~\ref{fig:figure-5}B, right). The joint marginal distribution of errors is strongly diagonal, which is expected based on the value for $\rho$ (Figure~\ref{fig:figure-5}B, left). The high correlation in errors leads to a speed up of 4-5 for 1 log unit selectivity optimization and 30-40x for 2 log unit selectivity optimization (Figure~\ref{fig:figure-5}B, right), despite the much higher per target errors.
\todo[inline]{add sigma selectivity}

@@

%
%
%  Discussion and Conclusions 
%
%
\section{Discussion and Conclusions}

@@

>>>>

\paragraph{Free energy calculation errors are correlated, accelerating selectivity optimization}
To quantify the correlation coefficient ($
\rho$) of the errors in our calculations, we built a Bayesian graphical model, as described in the methods section. Briefly, we modeled the absolute free energy ($G$) of each ligand in each phase (complex and solvent) as in equation~\ref{eq4}. The model was chained to the FEP+ calculations by providing the $\Delta G^{calc}_{phase,ij,target}$ as observed data, as in equation~\ref{eq6}. As in equation, the experimental data was modeled as a normal distribution centered around the true free energy of binding ($\Delta G^{true}_{i,target}$) corrupted by experimental error, which is assumed to be 0.3 kcal/mol from previous work done to quantify the uncertainty in publicly available data~\citep{BROWN2009420}. The reported IC50 values from each dataset were treated as data observations (Equation~\ref{eq12}) and the $\Delta G^{true}_{i,target}$ was assigned a weak normal prior (Equation~\ref{eq13}). The correlation coefficient was calculated for each sample according to equation~\ref{eq9}. 
The correlation coefficient $\rho$ for the CDK2/ERK2 calculations was quantified to be $0.5^{0.33}_{-0.23}$, indicating that the errors are largely uncorrelated between ERK2 and CDK2 (Figure~\ref{fig:figure-5}A, right). The joint marginal distribution of the error ($\epsilon$) for each target is symmetric, which is expected for cases in which $\rho$ is 0 (Supp. Figure~\ref{fig:sup-figure-2}). Despite the weak correlation in errors, the high per target accuracy of these calculations should have a 2-3x speed up for 1 log unit selectivity optimization, and a 20-30x speed up for 2 log unit selectivity optimization (Figure~\ref{fig:figure-5}A, right). 
The CDK2/CDK9 calculations show strong evidence of correlation, with a correlation coefficient of $0.70^{0.82}_{0.57}$ (Figure~\ref{fig:figure-5}B, right). The joint marginal distribution of errors is strongly diagonal, which is expected based on the value for $\rho$ (Figure~\ref{fig:figure-5}B, left). The high correlation in errors leads to a speed up of 4-5 for 1 log unit selectivity optimization and 30-40x for 2 log unit selectivity optimization (Figure~\ref{fig:figure-5}B, right), despite the much higher per target errors.
Quantifying $\rho$ for these calculations enables estimation of $\sigma_{selectivity}$, which is useful for estimating expected error for prospective studies, where the experimental values for $\Delta \Delta G_{selectivity}$ are not yet known. Based on the distribution quantified for $\rho$, the expected $\sigma_{selectivity}$ for the CDK2/CDK9 calculations is between 0.76 and 1.16 kcal/mol (Supp. Figure~\ref{fig:sup-figure-3}), which is in good agreement with the bootstrapped RMSE (Figure~\ref{fig:figure-4}, bottom). For the CDK2/ERK2 calculations, $\sigma_{selectivity}$ is expected to fall between 0.82 kcal/mol and 1.10 kcal/mol (Supp. Figure~\ref{fig:sup-figure-3}), which is also in good agreement with the bootstrapped RMSE (Figure~\ref{fig:figure-4}, top).

@@

%
%
%  Discussion and Conclusions 
%
%
\section{Discussion and Conclusions}
We have demonstrated, using a simple numerical model, the impact that free energy calculations with even weakly correlated errors can have on speeding up the optimization of selectivity in small molecule kinase inhibitors. While the expected speed up is dependent on the per target error of the method ($\sigma_{target}$), the speedup is also highly dependent on the correlation of errors made for both targets. Unsurprisingly, free energy methods have greater impact as the threshold for selectivity optimization goes from 10x to 100x. While 100x selectivity optimization is difficult to achieve, the expected benefit from free energy calculations is also quite high, with 1 and 2 order of magnitude speedups possible. 
To quantify the correlation of errors in two example systems, we gathered experimental data for two congeneric ligand series with experimental data for CDK2 and ERK2, as well as CDK2 and CDK9. These datasets, which had crystal structures for both targets with the same ligand co-crystallized, are exemplify the difficulty in predicting selectivity. The dynamic range of selectivity for both systems is incredibly narrow, with most of the perturbations not having a major impact on the overall selectivity achieved. Furhter, the data was reported with unrelaible experimental uncertainties, which makes quantifying the errors made by the free energy calculations difficult. This issue is common when considering selectivity, as many kinase-oriented high throughput screens are carried out at a single concentration and not highly quantitative. Work is being done to increase the availability of publicly available, high quality biophysical data for selectivity measurements, which will benefit future work on predicting selectivity using physical models and machine learning techniques. 
\todo[inline]{insert a discussion of outliers in the FEP+ calculations here?}
\todo[inline]{discuss quantification of rho (is there a way to know a priori what rho might be?)}
\todo[inline]{discussion of the expected speedup and sigma based on the quantification of rho}
\todo[inline]{extensions include: separating statistical error (unless this gets added into the results section}
\todo[inline]{Do we want to mention protein mutation FEP+ to enable cycle closure analysis for certain highly related targets?}
%
\newpage
%

@@

\begin{figure}[h]
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/supp_figure3.pdf}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:sup-figure-3}
{\bf Correlation reduces the expected error for selectivity predictions} \\
As corelation coefficient $\rho$ increases, $\sigma_{selectivity}$ decreases. The intersection between CDK2/CDK9 $\sigma_{selectivity}$ (green curve) and $\rho$ (black distribution) indicates the range of expected $\sigma_{selectivity}$ values. The intersection for CDK2/ERK $\sigma_{selectivity}$ (blue curve) and $\rho$ (gray distribution) suggests the expected $\sigma_{selectivity}$ range for that set of calculations. 
}
\end{fullwidth}
\end{figure}
